---
title: "Structural Bioinfo Class11"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


##PDB database composition statistics
Download PDB CSV Statistiscs from http://www.rcsb.org/stats/summary

```{r}
pdbstats <- read.csv("Data Export Summary.csv", row.names = 1)
pdbstats
```

Edit table format
```{r}
library(knitr)
kable(pdbstats)
```

Determine percentage of structures by X-ray & Electron Microscopy
```{r}
#Determine TOTAL structures
nstruc <- sum(pdbstats$Total)

#Determine % structures by experimental method
EMpro <- pdbstats$Total / nstruc
per <- round(EMpro, 4)*100

```

There are `r per[1]`% X-ray structures and `r per[3]`% Electron Microscopy structures in the PDB database as of `r Sys.Date()`
```{r}
#add percent to table
nstats <- pdbstats
nstats$Percent <- per
kable(nstats)
```

Proportion of structures are protein?
```{r}
proteinpro <- round(sum(nstats$Proteins) / nstruc *100, 2)
```
`r proteinpro`% of structures that are protein as of `r Sys.Date()`

There are 1130 HIV-1 protease structures in current PDB.

##Bio3D in R
load package with **library()** function
```{r}
library(bio3d)
```

```{r}
pdb <- read.pdb("1hsg")
pdb
```
198 residues. H2O and MK1 non-protein residues

find attributes
```{r}
attributes(pdb)
```

PDB$atom is a dataframe R object
```{r}
str(pdb$atom)
```

```{r}
#print xyz data
pdb$xyz

#examine table dimensions
dim(pdb$xyz)
```

##Atom Selection
Select C-alpha atoms
```{r}
ca.inds <- atom.select(pdb, "calpha")
ca.inds

```

Select few C-alpha atoms
```{r}
head(pdb$atom[ca.inds$atom,])
```

Select xyz coordinates
```{r}
head(pdb$xyz[, ca.inds$xyz])
```

Select Chain A
```{r}
a.inds <- atom.select(pdb, chain="A")

```

##working with multiple PDB files
```{r}
#download some example files
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <- get.pdb(ids, split = TRUE)

```
```{r}
# Extract and align the chains we are interested in
pdbs <- pdbaln(files, fit = TRUE)

# Print to screen a summary of the 'pdbs' object
pdbs
```

##PCA
```{r}
# find invariant core
core <- core.find(pdbs)

# superimpose all structures to core
pdbs$xyz = pdbfit(pdbs, core)

# Perform PCA
pc.xray <- pca(pdbs)

# plot PCA
plot(pc.xray)
```


```{r}
# Visualize first principal component
pc1 <- mktrj(pc.xray, pc=1, file="pc_1.pdb")

```



