---
title: "class09"
author: "Alyssa Shan"
date: "October 30, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Section 1

##Preparing data
```{r}
url <- "https://bioboot.github.io/bimm143_W18/class-material/WisconsinCancer.csv"
wisc.df <- read.csv(url)
```
##Editing input data
```{r}
#exlude data to columns 3-32
wisc.data <- as.matrix(wisc.df [,3:32])

#set the row names of wisc.data
row.names(wisc.data) <- wisc.df$id

```
##Create Diagnosis vector
```{r}
diagnosis <-  as.numeric(wisc.df$diagnosis == "M")
diagnosis
table(diagnosis)
```

##Exploratory Data Analysis

```{r}
#Q1. How many observations are in this dataset?
dim(wisc.data)

```
There are 30 observations in dataset

```{r}
##Q2. How many variables/features in the data are suffixed with _mean?
x <- grep("_mean", colnames(wisc.data))
length(x)
```
There are 10 features suffized with "_mean" in dataset

```{r}
##Q3. How many of the observations have a malignant diagnosis?
#Need to reference wisc.df since wisc.data has diagnosis column removed
y <- as.numeric(wisc.df$diagnosis == "M")
sum(y)

```
There are 212 malignant diagnosis

#Section 2
##Performing PCA
```{r}
#Check column means and stdev
colMeans(wisc.data)
apply(wisc.data, 2, sd)

#Perform PCA on wisc.data 
wisc.pr <- prcomp(wisc.data, scale=T)

#summarize PCA
summary(wisc.pr)
```

Q4. From your results, what proportion of the original variance is captured by the first principal components (PC1)?


Q5. How many principal components (PCs) are required to describe at least 70% of the original variance in the data?


Q6. How many principal components (PCs) are required to describe at least 90% of the original variance in the data?

##Interpreting PCA results
```{r}
#creat biplot of wisc.pr
biplot(wisc.pr)
attributes(wisc.pr)

#scatter plot PC1 vs PC2
plot(wisc.pr$x[,1], wisc.pr$x[,2],  col= diagnosis+1, xlab = "PC1", ylab = "PC2")

#scatter plot PC1 vs PC3
plot(wisc.pr$x[,c(1,3)], col = diagnosis+1, xlab = "PC1", ylab = "PC3")
```

##Variance explained
```{r}
#calculate variance
pr.var <- wisc.pr$sdev^2

#principal component: pve
pve <- pr.var/sum(pr.var)
pve

#plot variance 
plot(pve, xlab = "Principal Component", ylab = "Proportion of Variance Explained", ylim = c(0,1), type = "o")

#side by side plot
par(mfrow = c(1,2))

#Barplot variance
barplot(pve, ylab = "Percent of Variance Explained", names.arg = paste("PC", 1:length(pve)), las=2, axes = F)
#turn off standard axis measurements
axis(2, at=pve, labels=round(pve,2)*100)

#cumulative proportion of variance
plot(cumsum(pve), xlab = "Principal Component", 
     ylab = "Cumulative Proportion of Variance Explained", 
     ylim = c(0, 1), type = "o")

```

##Communicating PCA results
Q9. For the first principal component, what is the component of the loading vector (i.e.  wisc.pr$rotation[,1]) for the feature concave.points_mean?


Q10. What is the minimum number of principal components required to explain 80% of the variance of the data?
4 PCs

#Section 3
##Hierarchical clustering of case data
```{r}
# Scale the wisc.data 
data.scaled <- scale(wisc.data)

#Calculate distances
data.dist <- dist(data.scaled)

#create hclust
wisc.hclust <- hclust(data.dist, method = "complete")
plot(wisc.hclust)

```
Q11. Using the plot() function, what is the height at which the clustering model has 4 clusters?
20

##Selecting number of clusters
```{r}
#cut tree
wisc.hclust.clusters <- cutree(wisc.hclust, k=4)

table(wisc.hclust.clusters, diagnosis)
```

cluster 1 largely corresponds to malignant cells (with  diagnosis values of 1) whilst cluster 3 largely corresponds to benign cells (with diagnosis values of 0).

Q12. Can you find a better cluster vs diagnoses match with by cutting into a different number of clusters between 2 and 10?
```{r}
wisc.hclust.clusters <- cutree(wisc.hclust, k=2)

table(wisc.hclust.clusters, diagnosis)

wisc.hclust.clusters <- cutree(wisc.hclust, k=10)

table(wisc.hclust.clusters, diagnosis)
```

#Section 5
##Clustering on PCA results
```{r}
#create distance vector of 1st 7 PCs for clustering
d.pr <-  dist(wisc.pr$x[,1:7])

#create hclust
wisc.pr.hclust <- hclust(d.pr, method= "complete")
plot(wisc.pr.hclust)

#cut hierarchical clustering
wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=4)
table(wisc.pr.hclust.clusters, diagnosis)
```

#BONUS
## Predictive Modeling with PCA components
```{r}
url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)

plot(wisc.pr$x[,1:2], col=diagnosis+1)
points(npc[,1], npc[,2], col="blue", pch=16, cex= 2)

```

predicting potential malignancy of new samples. In this case, 1 sample might be maligant; while other sample, benign.