---
title: "Class12B: Structural Bioinformatics II"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Get HIV-Pr from PDB database and inspect
```{r}
library(bio3d)
file.name <- get.pdb("1hsg")

hiv <- read.pdb(file.name)

hiv
```

**Q1:What is the name of the two non protein resid values in this structure?**
water and MK1 = non-protein residues

## Prepare protein and ligand input files
```{r}
prot <- trim.pdb(hiv, "protein")
lig <- trim.pdb(hiv, "ligand")

prot
lig

write.pdb(prot, file = "1hsg_protein.pdb")
write.pdb(lig, file = "1hsg_ligand.pdb")


```

## Protein Docking using AutoDock Tools
**Q2: Can you locate the binding site visually? Note that
crystal structures normally lack hydrogen atoms, why?**


**Q3: Look at the charges. Does it make sense (e.g. based on your knowledge of the
physiochemical properties of amino acids)?**

## Docking with Vina
run this command: `"\Program Files (x86)\The Scripps Researc
h Institute\Vina\vina.exe" --config config.
text --log log.txt`

## Inspecting docking results
convert all.pdbqt to pdb format
```{r}
library(bio3d)
res <- read.pdb("all.pdbqt.txt", multi=T)
res

write.pdb(res, "results.pbd")
```

## Qualitatively determine docking
```{r}
ori <- read.pdb("ligand.pdbqt.txt")
rmsd(ori, res)
```



